---
title: "Assignment1_Mar13"
author: "Garrett Bullivant"
date: "2023-03-13"
output: html_document
---

	Is the visualization pleasing to look at? 
	Does the visualization accurately represent data?
	Can we understand what message the maker of the visualization is attempting to convey?
	Consider factors such as colour, size, use of images. Is this a ‘good’ data visualization? Why or why not?

Life expectancy 2007
- Challenging to see differences between countries because of 3D shape
- Uneccessary papyrus background and font
- Similar colours to denote different groups
- Redundancy with label and legend

Montrous costs 
- Character distracts from data
- Editorialized visualization
- Angle make it challenging to interpret 
- Lack of axes labels 

Healy, 2018
- Lack of any meaningful labels/titles 
- Interesting to see difference in interprtation between same data with different aspect ratios

2012 Presidential Run
- Source: opinions
- Data adds up to >100%?
- Colours are jarring

Wind map 
- Good contrast
- The motion is neat

```{r}
# load relevant libraries 
library(tidyverse)
library(ggplot2)
library(socviz)
library(gapminder)
```

```{r}
# make object

p <- ggplot(data = gapminder,
            mapping = aes(x=gdpPercap, y = lifeExp))
p + geom_point()

```
	Is the visualization pleasing to look at? 
	Does the visualization accurately represent data?
	Can we understand what message the maker of the visualization is attempting to convey?
	Consider factors such as colour, size, use of images. Is this a ‘good’ data visualization? Why or why not?

- aesthetics are not super appealing, kind of boring
- hard to see each individual point - a lot of overlap
- need units for each axes
- mostly legible
