---
title: "Assignment1_Mar18"
author: "Garrett Bullivant"
date: "2023-03-18"
output:
  pdf_document: default
  html_document: default
---


```{r}
# observe our dataset
library(gapminder)

gapminder
```
```{r}
# load relevant libraries 
library(tidyverse)
library(ggplot2)
library(socviz)
library(gapminder)
```

```{r}
# try plotting our gdp data by year
p <- ggplot(data = gapminder, 
            mapping = aes(x = year, y = gdpPercap))

p + geom_line()
```
```{r}
# try grouping our data first by country

p + geom_line(aes(group = country))
```

```{r}
# try breaking up our data according to contintents by using the facet
p + geom_line(aes(group = country)) + facet_wrap(~continent)
```
```{r}
# improving our plot
ggplot(data=gapminder,mapping=aes(x=year, y=gdpPercap))

# adding gray colour to lines, adding trend line, make y axis logarithmic, change scale to dollars, add axes labels and titles
p + geom_line(color='gray70', aes(group = country)) +
    geom_smooth(size=1.1,method='loess',se=FALSE) +
    scale_y_log10(labels=scales::dollar) +
  facet_wrap(~continent,ncol=5) +
  labs(x = 'Year',
    y = 'GDP per capita',
    Title = 'GDP per capita on Five Continents')
```
```{r}
# using facet grid to split data by 2 variables
p <- ggplot(data=gss_sm, mapping = aes(x=age, y=childs))
p + geom_point(alpha=0.2) + geom_smooth() + facet_grid(sex~race)
```

```{r}
# exploring stat functions - default stat is count
p <- ggplot(data=gss_sm, mapping = aes(x = bigregion))
p + geom_bar()
```
```{r}
# try using another stat - proportion 
p + geom_bar(mapping = aes(y = ..prop..))
```
```{r}
# need to create dummy group for grouping. 
p + geom_bar(mapping = aes(y = ..prop.., group = 1))

```
```{r}
# frequency plots
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar()
```

```{r}
# want to compare proportions across groups
p + geom_bar(position = 'fill')
```
```{r}
# now we can compare proportions across groups, but we can't see relative size respect to overall total
# faceting allows us to see proportional breakdown by region
p <- ggplot(data = gss_sm, mapping = aes(x = religion))

p + geom_bar(position = 'dodge', mapping = aes(y = ..prop.., group =
bigregion)) + facet_wrap(~bigregion, ncol = 1)
```
```{r}
# explore histograms
p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram()
```
```{r}
# we can also manually set the number of bins
p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram(bins = 10)
```
```{r}
# subset our data by using a character vector of 2 states 
oh_wi <- c("OH", "WI")

p <- ggplot(data = subset(midwest, subset = state %in% oh_wi),
mapping = aes(x = percollege, fill = state))

p + geom_histogram(alpha = 0.4, bins = 20)
```
```{r}
# using a density plot for continous variables
p <- ggplot(data = midwest,mapping = aes(x = area))
p + geom_density()
```
```{r}
# subset the data for our density plot
p <- ggplot(data = subset(midwest, subset = state %in% oh_wi), 
            mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3)
```



- What information can we learn from this visualization?
- Is this an example of objective, neutral data visualization? Why or why not?

- second visual uses less emotional language and is just static and still
- second visual is easier to interpret objectively 
- first visual is not neutral, but is any data visualization neutral? No. 
- Data visualizations can be factual without being neutral

```{r}
# explore organ dataset
organdata

# we can pipe data
organdata %>% select(1:6) |> sample_n(size = 10)
```

```{r}
# recall groups and facets. we can look at number of donors by years by country
p <- ggplot(data = organdata, mapping = aes(x = year, y = donors))

p + geom_line(aes(group = country)) + facet_wrap(~country)
```
```{r}
# make a boxplot
p <- ggplot(data = organdata, mapping = aes(x = country , y = donors))
p + geom_boxplot()

```
```{r}
# improve our boxplot with coord_flip
p + geom_boxplot() + coord_flip()
```
```{r}
# improve our plot further by listing the countries from high to low donor
p <- ggplot(data = organdata, mapping = aes(x = reorder(country, 
                                                        donors, na.rm = TRUE ), y = donors))
```

```{r}
# if we have fewer data observations, it can be helpful to show individual observations 
p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, 
                                                        na.rm=TRUE),y = donors, color = world))
p + geom_point() + coord_flip()
```

```{r}
# we can use geom_jitter o randomly and slightly nudge each observation for better visualization 
p + geom_jitter() + coord_flip()
```
```{r}
# processing data with a pipeline
by_country <- organdata |> 
  group_by(consent_law, country) |>
  summarize_if(is.numeric, funs(mean, sd), na.rm = TRUE) |>
  ungroup()
```

```{r}
# make a cleveland dot plot for categorical data 
p <- ggplot(data = by_country, mapping = aes(x = donors_mean,
                                             y = reorder(country, donors_mean), color = consent_law))

p + geom_point(size=3) + labs(x = 'Donor Procurement Rate', y = '', color = 'Consent Law') + theme(legend.position='top')
```
```{r}
# plotting text
# adding data labels; we can add text to our plots with geom_text
p <- ggplot(data = by_country, mapping = aes(x = roads_mean, 
                                             y = donors_mean))

p + geom_point() + geom_text(mapping = aes(label = country))
```
```{r}
# we can adjust the position of our text
p + geom_point() + geom_text(mapping = aes(label = country), hjust = 0)
```
```{r}
# for more flexible text formatting, we can use ggrepel 
# load libraries
# install.packages('ggrepel')
library(ggrepel)

# we will use a new dataset about historical united states presidents
p <- ggplot(elections_historic, aes(x = popular_pct, y = ec_pct, label = winner_label))
p + geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +
geom_point() +
geom_text_repel() +
scale_x_continuous(labels = scales::percent) +
scale_y_continuous(labels = scales::percent) +
labs(x = "Winner’s share of Popular Vote", y = "Winner’s share of Electoral
College Votes", title = "Presidential Elections: Popular & Electoral
College Margins", subtitle = "1824 2016")
```

