---
title: "Assignment1_Apr1"
author: "Garrett Bullivant"
date: "2023-04-01"
output: html_document
---

Activity: Comparing data visualizations
How are they different? What does each visualization ‘do’?
What are the pros and cons of each?
- one is dynamic visualization, one is statiic
- second visualization allows the user to interrogate 
- second visualization is visually appealing
- the word needs to be available to search up for second visualization
- con of first visualization is some scientific notation that may not be accessible

Activity: Types of changes
- identity preserving type of change
- spatial position changes and retinal elements stay the same
- add other gender
- maybe include less disciplines?

```{r}
#install.packages('rAmCharts')
library(rAmCharts)
amPlot(iris, col = c('Sepal.Length', 'Sepal.Width'), 
type = c("line", "step"),
zoom = TRUE, legend = TRUE)
```

```{r}
# install packages neccessary for dynamic ggplot

#install.packages('gganimate')
#install.packages('gifski')

library(ggplot2)
library(gganimate)
library(gifski)
library(gapminder)

# first we make a static plot
p <- ggplot(gapminder, aes(x = gdpPercap, y=lifeExp, size = pop,
colour = country)) +
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(x = "GDP per capita", y = "Life expectancy")
p

```
```{r}
# lets animate plot for changes over time
p2 <- p + transition_time(year) +
  labs(title = "Year: {frame_time}")
animate(p2, duration = 10, renderer = gifski_renderer())

```

SO COOL!!!!

```{r}
install.packages('shiny')
library(shiny)
```

