---
title: "Assignment1_Mar25"
author: "Garrett Bullivant"
date: "2023-03-25"
output: html_document
---

```{r}
# load libraries
library(ggplot2)
library(tidyverse)
library(socviz)
library(gapminder)
library(ggrepel)

# build colour palette
model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors

# 
p0 <- ggplot(data = gapminder, mapping = aes(x = log(gdpPercap), y =
lifeExp)) 

p1 <- p0 + geom_point(alpha = 0.2) + 
  geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) + 
  geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3), 
              aes(color = "Cubic Spline", fill = "Cubic Spline")) + 
  geom_smooth(method = "loess", aes(color = "LOESS", fill = "LOESS"))

# add a legend
p1 + scale_color_manual(name = "Models", values = model_colors) +
  scale_fill_manual(name = "Models", values = model_colors) + 
  theme(legend.position = "top")              

# explore data
gapminder
str(gapminder)

# make our model object
out <- lm(formula = lifeExp ~ gdpPercap + pop + continent, data = gapminder)

# explore model
str(out)
summary(out)
```
```{r}
# make dataframe for predictions
min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)

pred_df <- expand.grid(gdpPercap = (seq(from = min_gdp, to = max_gdp, length.out = 100)),
                       pop = med_pop,
                       continent = c("Africa", "Americas", "Asia", "Europe", "Oceania"))

# make predictions
pred_out <- predict(object = out, newdata = pred_df, interval = "predict")
head(pred_out)

```
```{r}
#bind predictions and data
pred_df <- cbind(pred_df, pred_out)

head(pred_df)
```
```{r}
p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")), 
            aes(x = gdpPercap, y = fit, ymin = lwr, ymax = upr , color = continent, fill = 
                  continent, group = continent))

p + geom_point(data = subset(gapminder, continent %in% c("Europe", "Africa")), 
               aes(x = gdpPercap, y = lifeExp, color = continent), 
               alpha = 0.5, 
               inherit.aes = FALSE) +
  geom_line() + 
  geom_ribbon(alpha = 0.2, color = FALSE) + 
  scale_x_log10(labels = scales::dollar)
```

Activity 

Research question
- How is life expectancy influenced by the population of a given country?



```{r}
p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")), 
            aes(x = pop, y = fit, ymin = lwr, ymax = upr , color = continent, fill = 
                  continent, group = continent))

p + geom_point(data = subset(gapminder, continent %in% c("Europe", "Africa")), 
               aes(x = pop, y = lifeExp, color = continent), 
               alpha = 0.5, 
               inherit.aes = FALSE) +
  geom_line() + 
  geom_ribbon(alpha = 0.2, color = FALSE) 
```

```{r}
# load broom library
install.packages("broom")
library(broom)

# use tidy to gather component level stats
out_conf <- tidy(out)
out_conf |> round_df()


```

```{r}
# plot component level stats
p <- ggplot(out_conf, mapping = aes(x = term, y = estimate))
p + geom_point() + coord_flip()
```

```{r}
# use confint to produce confidence intervals
out_conf <- tidy(out, conf.int = TRUE)
out_conf %>% round_df()
```

```{r}
# clean up our visualization with tidy
out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")

# include confidence intervals
p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate), y = estimate, ymin = conf.low, ymax = conf.high))
p + geom_pointrange() + coord_flip() + labs(x = "", y = "OLS Estimate")
```
```{r}
# augment adds observation level statistics
out_aug <- augment(out)
head(out_aug) |> round_df()

```
```{r}
# we can now plot observation level stats - residuals vs fitted values
p <- ggplot(data = out_aug, mapping = aes(x = .fitted, y = .resid))
p + geom_point()
```
```{r}
# finally we can use glance to gather model level statistics 

glance(out) |> round_df()
```
```{r}
# Using broom for grpuped analysis
eu77 <- gapminder |> filter(continent == "Europe", year == 1977)
fit <- lm(lifeExp ~ log(gdpPercap), data = eu77)
summary(fit)
```
```{r}
# nesting data
out_le <- gapminder |>
  group_by(continent, year) |>
  nest()

out_le
```

```{r}
# we can now easily pick out data by continent and year
out_le |> filter(continent == "Europe" & year == 1977) |> 
  unnest()
```
```{r}
# create a function that fits our model to a dataframe
fit_ols <- function(df) {lm(lifeExp ~ log(gdpPercap), data = df)}

# apply model to each row
out_le <- gapminder |>
  group_by(continent, year) |>
  nest() |> 
  mutate(model = map(data, fit_ols))
out_le
```

```{r}
# tidy up our data
out_tidy <- gapminder |> group_by(continent, year) |> nest() |>
  mutate(model = map(data, fit_ols), tidied = map(model, tidy)) |>
  unnest(tidied) |> filter(term %nin% "(Intercept)" & continent %nin% "Oceania")
```

```{r}
p <- ggplot(data = out_tidy, mapping = aes(x = year, y = estimate, ymin = estimate - 2*std.error, ymax = estimate + 2*std.error, group = continent, color = continent))

p + geom_pointrange(position = position_dodge(width = 1)) + 
  scale_x_continuous(breaks = unique(gapminder$year)) + 
  theme(legend.position = "top") +
  labs(x = "Year", y = "Estimate", color = "Continent")
```


Activity
How did this diagram of the Brooks use rational, moral, and emotional appeal to make a case to its audiences?

Using the diagrams to show pictures of people on the boat shows how horrible the conditions are instead of just using people/boats. It uses emotional persuasian to pull on empathy. Furthermore, I think this also uses moral appeal - it appeals to the audiences moral values. 